---
title: "Regression Analysis Part 3"
author: "Cody"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Fit the multiple regression model, using two of your categorical predictors and 2-3 numerical predicitors, and include all pairwise interaction terms

**Initial linear model summary:**

```{r, echo=FALSE, comment=''}
setwd("/Users/crorick/Documents/MS\ Applied\ Stats\ Fall\ 2023/MA4710/Course\ Project")
set.seed(10)

#Load in NYC Airbnb data
NYC <- read.csv('AB_NYC_2019.csv')
NYC <- subset(NYC, select = -c(latitude,longitude, price, calculated_host_listings_count, last_review, host_id, neighbourhood, availability_365, host_name, name, id))
NYC <- na.omit(NYC)

#Make reviews per month reviews per year instead
NYC$reviews_per_year <- NYC$reviews_per_month * 12

#Reducing the data points down to a random sample of 1000 for better and
# faster processing
NYC <- NYC[sample(nrow(NYC), 1000), ]

#There is not enough data on staten Island so removing it from the data
NYC <- NYC[!(NYC$neighbourhood_group == 'Staten Island'),]

#make sure factor variables are in data as factors
NYC$neighbourhood_group <- as.factor(NYC$neighbourhood_group)
NYC$room_type <- as.factor(NYC$room_type)

NYC.lm <- lm(reviews_per_year  ~  (minimum_nights + number_of_reviews + neighbourhood_group + room_type)^2, NYC)

summary(NYC.lm)
```

## 2 Assess whether or not any of the interaction terms should be included in the model by investigating the p-values associated with each interaction term

**Final linear model summary:**

```{r, echo=FALSE, comment=''}
reduced.NYC.lm1 <- update(NYC.lm, ~ . - neighbourhood_group:number_of_reviews)
reduced.NYC.lm2 <- update(reduced.NYC.lm1, ~ . - neighbourhood_group:minimum_nights)
reduced.NYC.lm3 <- update(reduced.NYC.lm2, ~ . - number_of_reviews:room_type)
reduced.NYC.lm4 <- update(reduced.NYC.lm3, ~ . - number_of_reviews:minimum_nights)
summary(reduced.NYC.lm4)
```

## 3 Create a set of diagnostic plots and interpret them

**How does the linearity assumption appear to be met?**

```{r, echo=FALSE, comment=''}
plot(reduced.NYC.lm4, which = 1)
```

* The linearity assumption does not appear to be a perfect assumption as the residuals are not evenly scattered around the fitted line.

**How does the normality assumption appear to be met?**

```{r, echo=FALSE, comment=''}
plot(reduced.NYC.lm4, which = 2)
```

* The data skews towards extremes at the higher end of the theoretical quantiles, normality is not met.

**How does the homoscedasticity assumption appear to be met?**

```{r, echo=FALSE, comment=''}
plot(reduced.NYC.lm4, which = 3)
```

* Homoscedasticity does not appear to be met as the scale-location plot trend line increases with increasing fitted values, which indicates a increasingly large variance with increasingly large predictions.

**Are there any influential observations in your model? How do you know they are influential?**

```{r, echo=FALSE, comment=''}
plot(reduced.NYC.lm4, which = 5)
```

* I consider Point 39876 an influential point because the cook's distance is greater than 0.5, there are no other points with a cook's distance greater than this.

## 4 Interpret the presence of any interaction terms in the model:

**If there is an interaction between a numerical predictor and a categorical predictor, provide a verbal interpretation of the associated regression coefficients.**

* There is an interaction between the room type and the minimum nights of stay. For a stay in a private room you, if you increase the minimum nights required for a stay there is a positive impact on the review frequency. This association is less clear for shared rooms.

**If there is an interaction between two categorical predictors, provide verbal interpretations of the associated regression coefficients.**

* There is an interaction between neighborhood group and room type. The burrough the Airbnb is located in combined with whether the room is shared or private has an impact on the frequency of reviews. A shared room in Manhattan for example will raise the review frequency more than a private room in Manhattan. And a private room in Brooklyn will raise the review frequency more than a private room in Manhattan. 

## 5 Are there any categories in one of your categorical variables which could be combined together? 

**Perform a set of simultaneous hypothesis tests associated with the different levels of one of your categorical variables (the multcomp or car package will be very useful in this regard).**

```{r, echo=FALSE, comment=''}
library(multcomp)
diff.glht <- glht(reduced.NYC.lm4, 'neighbourhood_groupManhattan - neighbourhood_groupBrooklyn = 0')
summary(diff.glht)
```

* Combining burroughs Manhattan and Brooklyn together does not appear to have a significant impact on the model fit.

**If you get insignificant p-values associated with combining two (or more) categories together, refit your model and perform diagnostics to see what assumptions are still met (or not met)**

```{r, echo=FALSE, comment=''}
new_neighbourhood_group <- NYC$neighbourhood_group
levels(new_neighbourhood_group)[2:3] <- "Brooklyn.Manhattan"
reduced.NYC.lm5 <- update(reduced.NYC.lm4, ~ . -neighbourhood_group - neighbourhood_group:room_type + new_neighbourhood_group + new_neighbourhood_group:room_type)
anova(reduced.NYC.lm4, reduced.NYC.lm5)
cat("adjusted r squared for old model:\n")
cat(summary(reduced.NYC.lm4)$adj.r.squared)
cat("adjusted r squared for new model with brooklyn and manhattan combined:\n")
cat(summary(reduced.NYC.lm5)$adj.r.squared)
```

* This combination increases the correlation, as the adjusted R squared value increases in the updated model that combines Brooklyn and Manhattan.
* There is not a significant difference between the old model and the newly refit model,  showing these variables could be combined.  